Make any stupid plot you want

### 2021-02-10 draft/sections/intro.tex
contribution + intro section

\subsection{what is the problem we're trying to solve}
% build better viz libaries
%formally define what constraints on libraries are
% make something seperable and scalable

\subsection{what are we doing}
constraints: toplogy of dataset + structure on the variables
implementation: be fully feature
we can haz compose the things into larger structure. \& has framwework to do larger things w/ algebraic formalizations.

\subsection{What even is a data visualization library?}
This work aims to develop a model of visualization such that a tool built using this model could support visualizations as varied as those of Du Bois in figure~\ref{fig:}; to do so, we first discuss the criteria by which a visualization is evaluated.

Byrne et al. describe visualizations as having graphic and figurative components; graphic representations are maps from data values to visual encodings\cite{bertinSemiologyGraphicsDiagrams2011a}, while figurative representations have meaning due to their similarity in shape to external concepts \cite{byrneAcquiredCodesMeaning2016}. In figure~\ref{fig:intro_dbc}, Du Bois combines a graphical representation where glyph size varies by population with a figurative representation of those glyphs as the countries the data is from, which means that the semantic nominal and numerical properties of the data are preserved in the graph. Norman states that when the properties of the representation match the properties of the information being represented\cite{norman_things_smart, NaturalnessPrincipleInfoVis} then the visualizations are more understandable 
, but does not give criteria for what it means for the properties to match. Tufte specifies that visual representations must be in proportion to the quantitative data being represented, but otherwise his notion of graphic integrity is heavily context dependent\cite{tufteVisualDisplayQuantitative2001}.

Mackinlay

@@

\subsection{contribution}

\begin{enumerate}
    \item fiber bundle representation of data and graphics
    \item formalization of variable type structure preservation as monoid action equivariant maps between fibers 
    \item formalization of continuity preservation as equivariant map between base space
    \item algebraic sum for composition of graphics
    \item functional approach to force seperation of scope and concerns
    \item prototype  
\end{enumerate}

\end{document}

>>>>

In data visualization it is assumed the graphical representation of data match the properties of the data, and in this work we propose that the mathematical notion of equivariance formalizes this expectation. We meld the infoviz communities interest in heterogenous often discrete datasets with the scientific visualization communities emphasis on continuous and sometimes topologically complex datasets. To demonstrate the practical value of our model, we propose a model driven re-architecture of the artist layer of the Python visualization library Matplotlib. In addition to providing a way to ensure the library preserves structure, we propose a functional approach to  improve modularity, maintainability, and point to ways in which the library could better support concurrency and interactivity.

\subsection{Background}
This work aims to develop a model of visualization such that a tool built using this model could support visualizations as varied as those of Du Bois in figure~\ref{fig:intro_dubois}; to do so, we first discuss the criteria by which a visualization is evaluated. As described by Mackinlay, a visualization tool produces a graphical design and an image rendered based on that design. He defines the graphical design as the set encoding relations from data to visual representation, and the design appied on an idealized space is what throughout this paper we will refer to as graphics \cite{mackinlayAutomatingDesignGraphical1986}. Mackinlay  proposes that a visualization tool can be evaluated by its expressiveness and effectiveness; A visualization tool's expressiveness is a measure of how much of the structure of the data the tool encodes, while effectiveness describes how much design choices are made in deference to perceptual saliency \cite{clevelandResearchStatisticalGraphics1987,clevelandGraphicalPerceptionTheory1984,chambersGraphicalMethodsData1983a}. Mackinlay's definition of expressiveness is formalized at the variable level, for example that ordinal variables must map to ordinal positions. He uses a scatter chart for discrete data states that a line chart must have continuous data, but does not describe how to prove this. 
While Mackinlay is strictly focusing on the graphical components of a visualization, Byrne et al. propose that visualizations also have figurative representations that have meaning due to their similarity in shape to external concepts \cite{byrneAcquiredCodesMeaning2016}. In figure~\ref{fig:intro_dbc}, Du Bois combines a graphical representation where glyph size varies by population with a figurative representation of those glyphs as the countries the data is from, which means that the semantic nominal and numerical properties of the data are preserved in the graph. Tufte specifies that visual representations must be in proportion to the quantitative data being represented for a chart to be faithful and that there should be no extra information in the graphic or figurative elements of the graph, but otherwise his notion of graphic integrity is heavily context dependent\cite{tufteVisualDisplayQuantitative2001}. As is Norman's Naturalness Principal, which states that visualizations are more understandable  when the properties of the representation match the properties of the information being represented\cite{norman_things_smart}. Bertin takes it as a given that data properties match visual properties, so much so that Munzner argues it is inherently built into his classification system \cite{munznerVisualizationAnalysisDesign2014}. In this work, we propose that this property matching can be formally defined as equivariant maps from data to graphic representation.

@@

\subsection{contribution}
The contribition of this work is a model we call the topological artist model (TAM) in which data and graphics can be viewed as sections of fiber bundles. This model allows for (1) decomposing the translation of data fields (variables) into visual channels via an equivariant map on the fibers and (2) a topology-preserving map of the base spaces that translates the dataset connectivity into graphical elements. Furthermore, this model supports an algebraic sum operation such that more complex visualizations can be built from simple ones. To demonstrate the practical value of the model, we built a prototype where we represent the topological base spaces using triangulation, make use of programming types for the fiber, and build on Matplotlib's existing infrastructure for the rendering. 
\end{document}
